Drivers of sex differences in consumption

### 2017-06-16 sex-difference/sex-difference.Rmd
sex difference present

@@

```{r load_data}
year = 2011
bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
  mutate(is_female=sex=='female', is_white=race=='white')

@@

sex_icd = read_tsv('sex_codes.tsv', col_types=cols(code='c')) %>%
  rename(dx_code=code, dx_desc=short_desc)
```

```{r load_dx}
dx = read_tsv(sprintf('../../data/dx/dx_car_claims_%i.tsv', year), col_types=cols(diagnosis='c')) %>%
  select(bene_id=BENE_ID, dx_code=diagnosis) %>%
  left_join(sex_icd, by='dx_code')

@@

total_pde = pde %>% count(bene_id) %>% rename(n_claims=n)
```

@@

```{r}
sex_icd %>%
  left_join(dx_counts, by='dx_code') %>%
  kable
```

@@

It seems that females consume more antibiotics even within each age bracket (except for those over 95).

@@

```{r drugs_by_sex}
usage_by_sex = function(drug) {
  bene %>%
    left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id) %>% rename(n_claims=n), by='bene_id') %>%
    replace_na(list(n_claims=0)) %>%
    group_by(sex) %>%
    summarize(mean_n_claims=mean(n_claims)) %>%
    ungroup() %>%
    spread(sex, mean_n_claims) %>%
    mutate(drug=drug)
}

@@

```{r drug_by_sex_plot}
ubs %>%
  mutate(highlight=antibiotic %in% c('azithromycin', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'fosfomycin', 'nitrofurantoin')) %>%
  ggplot(aes(x=male, y=female, label=antibiotic, color=highlight)) +
  geom_text() +
  geom_point() +
  geom_abline() +
  xlab('claims per 1,000 male beneficiaries') +
  ylab('claims per 1,000 female beneficiaries') +
  theme_minimal() +
  coord_fixed()
```

For a more quantitative approach, this is a table showing, for each antibiotic, the CPKP for female and males, the difference (residual) between the two, and the residual fraction (the residual over the mean of the female and male consumptions). Positive residuals means more consumption by females.

```{r}
dbs %>% head %>% kable
```

```{r drug_by_sex_residuals, eval=FALSE}
dbs %>%
  select(antibiotic, sex, cpkp) %>%
  spread(sex, cpkp) %>%
  mutate(residual=female-male, residual_fraction=residual/(0.5 * (male + female))) %>%
  arrange(desc(abs(residual))) %>%
  head(10) %>%
  kable
```

@@

```{r azithromycin_usage}
na_to_zero = function(x) if_else(is.na(x), 0L, x)

@@

```{r azithromycin_usage_table}
azi_usage %>%
  group_by(age_group, sex) %>%
  summarize(n_beneficiaries=n(),
            mean_n_azithromycin_claims=mean(n_claims),
            med_n_carrier_dates=median(n_carrier_dates),
            med_n_outpatient_dates=median(n_outpatient_dates),
            med_n_inpatient_dates=median(n_inpatient_dates)) %>%
  kable
```

@@

```{r azithromycin_usage_model}
library(broom)
azi_usage %>%
  mutate(is_female=sex=='female') %>%
  glm(n_claims ~ age + is_female + n_carrier_dates + n_inpatient_dates + n_outpatient_dates, data=., family='poisson') %>%
  tidy %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100, 2)) %>%
  kable
```

@@

>>>>

# To do

- Separate beneficiaries into "high users" and "low users" based on the median number of outpatient claim dates
- Try segmenting azithromycin consumption by low/high in wo/men and by RC+/- in wo/men. We want to say that RC, more than low/high, accounts for differences in azithromycin risk, i.e., it's that women get more respiratory complaint diagnoses, not that women just have more complaints. Or vice versa.

@@

```{r load_data}
year = 2011

@@

sex_icd = read_tsv('sex_codes.tsv', col_types=cols(code='c')) %>%
  rename(dx_code=code, dx_desc=short_desc)

dx = read_tsv(sprintf('../../data/dx_car_claims_%i.tsv', year), col_types=cols(diagnosis='c')) %>%
  select(bene_id=BENE_ID, dx_code=diagnosis) %>%
  left_join(sex_icd, by='dx_code')

@@

bene_dx = dx %>%
  select(bene_id, diagnosis_type) %>%
  distinct() %>%
  mutate(val=TRUE) %>%
  spread(diagnosis_type, val)

total_pde = pde %>% count(bene_id) %>% rename(n_claims=n)

bene = read_tsv(sprintf('../bene_%i.tsv', year)) %>%
  mutate(is_female=sex=='female', is_white=race=='white') %>%
  left_join(total_pde, by='bene_id') %>% replace_na(list(n_claims=0)) %>%
  left_join(bene_dx, by='bene_id') %>% replace_na(list(acute_rc=FALSE, copd=FALSE, uti=FALSE))
```

@@

```{r}
sex_icd %>%
  left_join(dx_counts, by='dx_code') %>%
  arrange(diagnosis_type, desc(n)) %>%
  mutate(n=format(n, big.mark=',')) %>%
  kable(align='lllr')
```

@@

@@

```{r drugs_by_sex}
usage_by_sex = function(drug) {
  bene %>%
    select(-n_claims) %>%
    left_join(pde %>% filter(antibiotic==drug) %>% count(bene_id) %>% rename(n_claims=n), by='bene_id') %>%
    replace_na(list(n_claims=0)) %>%
    group_by(sex) %>%
    summarize(mean_n_claims=mean(n_claims)) %>%
    ungroup() %>%
    spread(sex, mean_n_claims) %>%
    mutate(drug=drug)
}

@@

```{r drug_by_sex_plot}
ubs %>%
  mutate(highlight=antibiotic %in% c('azithromycin', 'ciprofloxacin', 'trimethoprim/sulfamethoxazole', 'fosfomycin', 'nitrofurantoin')) %>%
  ggplot(aes(x=male, y=female, label=antibiotic, color=highlight)) +
  scale_color_manual(values=c('black', 'red')) +
  geom_text() +
  geom_point() +
  geom_abline() +
  xlab('claims per 1,000 male beneficiaries') +
  ylab('claims per 1,000 female beneficiaries') +
  theme_minimal() +
  coord_fixed()
```

Alternatively, I look at the mean usage (across men and women), the residual (the difference between the fraction of female beneficiaries taking the drug and male beneficiaries), and the fractional residual (the residual over the mean). I keep only those drugs with mean fraction at least 1%, and I also keep fosfomycin.

```{r}
ubs %>%
  mutate(mean_fraction=0.5*(male+female),
         residual=female-male,
         fractional_residual=residual/mean_fraction) %>%
  filter(mean_fraction > 0.01 | antibiotic=='fosfomycin') %>%
  ggplot(aes(x=mean_fraction, y=fractional_residual, label=antibiotic)) +
  geom_point() +
  geom_text()
```

This makes the same four look interesting; maybe levofloxacin and metronidazole would be next. Levo fits in with the ARTI story; metro does not.

## UTIs

Are the increased consumptions of cipro and TMP/SMX explained by UTIs? That is, do men and women who have UTIs get cipro and TMP/SMX at similar rates, and it's just that women get more UTIs than men?

### Cipro

```{r}
outcome_bene = pde %>% filter(antibiotic=='ciprofloxacin') %$% bene_id %>% unique
epi = bene %>%
  mutate(outcome=bene_id %in% outcome_bene) %>%
  count(sex, uti, outcome) %>%
  ungroup()
```

```{r}
group_by_all_ = function(df, group_name) {
  bind_rows(df, mutate_(df, .dots=setNames(list(~'all'), group_name))) %>%
    group_by_(group_name)
}

epi %>%
  group_by_all_('sex') %>%
  summarize(n_cases=sum(n)) %>%

```

@@

```{r azithromycin_usage, eval=FALSE}
na_to_zero = function(x) if_else(is.na(x), 0L, x)

@@

```{r azithromycin_usage_table, eval=FALSE}
azi_usage %>%
  group_by(age_group, sex) %>%
  summarize(n_beneficiaries=n(),
            mean_n_azithromycin_claims=mean(n_claims),
            med_n_carrier_dates=median(n_carrier_dates),
            med_n_outpatient_dates=median(n_outpatient_dates),
            med_n_inpatient_dates=median(n_inpatient_dates)) %>%
  kable
```

@@

```{r azithromycin_usage_model, eval=FALSE}
library(broom)
azi_usage %>%
  mutate(is_female=sex=='female') %>%
  glm(n_claims ~ age + is_female + n_carrier_dates + n_inpatient_dates + n_outpatient_dates, data=., family='poisson') %>%
  tidy %>%
  mutate(pct.increase=round((exp(estimate)-1.0)*100, 2)) %>%
  kable
```

@@

# Supplement

At every age, women consume more antibiotics than men.
